---
title: "Final Project"
author: "Cihan"
date: "14 05 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **R Final Project**

> The scope of this project is the enviromental analyze of Izmir from the s datasets generated by the Izmir Municipality (IBB Acik Veri
Portali https://acikveri.bizizmir.com/dataset). In this project we will try to analyze the Water Consumption and Production and The Amount of Food Distriubted for the Animals.

***

## Imports
```{r}
library(ggplot2)
library('stringr')
library(dplyr)

```

***

### Loading Dataset
```{r}
water.consumption <- read.csv("C:\\Users\\sakma\\Desktop\\R Files\\Final Project\\csv_files\\yillik-mahalle-su-tuketimi.csv", header = TRUE, encoding="UTF-8", stringsAsFactors=FALSE,sep=";")

animal.food <- read.csv("C:\\Users\\sakma\\Desktop\\R Files\\Final Project\\csv_files\\dagitilan-mama-miktari.csv", header = TRUE, encoding="UTF-8", stringsAsFactors=FALSE, sep=";")

water.production <- read.csv("C:\\Users\\sakma\\Desktop\\R Files\\Final Project\\csv_files\\su-uretimi-kaynak-dagilimi.csv", header = TRUE, encoding="UTF-8", stringsAsFactors=FALSE,sep=";")
```

***

Change the column names
```{r}
colnames(water.consumption) <- c("year","month","district","neighborhood","user_count","avg_consumption")
colnames(water.production) <- c("year","month","source","amount_m3")
colnames(animal.food) <- c("year","month","district","neighborhood","distributor","count","unit")


```

### **Data visualization and descriptive stats**

> Let's see some of data
```{r echo=FALSE}
paste("Water Consumption")
head(water.consumption, 10)
paste("Water Production")
head(water.production, 10)
paste("Animal Food")
head(animal.food, 10)
```

>- Summiraze all three data
```{r echo=FALSE}
paste("Summary of Water Consumption")
summary(water.consumption)
paste("Summary of Water Production")
summary(water.production)
paste("Summary of Animal Food")
summary(animal.food)

```


>There are some mistakes in the data.

1. In **Water Consumption**, *'year'* and *'month'* columns' type should be String, *'avg_consumpiton'* column kept as String and numbers are seperated by **','**. We'll replace these ',' with '.' and convert the type into **float**.

2. In **Water Production**, *'year'* and *'month'* columns' type should be String, *'amount_m3* should kept as **float**.

3. In **Animal Food**, *'year'* and *'month'* columns' type should be String.

```{r echo=TRUE}

water.consumption$avg_consumption <- str_replace_all(water.consumption$avg_consumption , ',', '.')

paste("Edit for Water Consumption")
water.consumption$avg_consumption = as.numeric(as.character(water.consumption$avg_consumption))
water.consumption$year = as.character(as.numeric(water.consumption$year))
water.consumption$month = as.character(as.numeric(water.consumption$month))
paste("Types of Features in Water Consumption after Edit")
sapply(water.consumption, mode)

paste("Edit for Water Production")
water.production$amount_m3 = as.numeric(as.character(water.production$amount_m3))
water.production$year = as.character(as.numeric(water.production$year))
water.production$month = as.character(as.numeric(water.production$month))
paste("Types of Features in Water Production after Edit")
sapply(water.production, mode)

paste("Edit for Animal Food")
animal.food$year = as.character(as.numeric(animal.food$year))
animal.food$month = as.character(as.numeric(animal.food$month))
paste("Types of Features in Water Animal Food after Edit")
sapply(animal.food, mode)


```



> Let's take the Water Consumption of 202

```{r}
consumption.2020 = subset(water.consumption, subset = water.consumption$year == 2020)
head(consumption.2020,10)
summary(consumption.2020)


#We can see the most consumer districts in 2020 in Izmir.
df <- consumption.2020 %>% 
  group_by(district) %>% 
  summarize(total_consumption = sum(avg_consumption))

df <- df %>%
  arrange(desc(total_consumption))

df %>% tbl_df %>% print(n=40)

p<-ggplot(data=df, aes(x=district, y=total_consumption)) +
  geom_bar(stat="identity")

p

```

